### 2019-10-20 referee_reply/referee_reply.tex
Adding *.swp to .gitignore, renaming Referee reply to referee_reply

\documentclass{article}
\title{Reply to referee report on \verbatim{JHEP_111P_0819}}
\author{Adarsh Pyarelal and Shufang Su}

\begin{document}
\maketitle

We would like to thank the referee for their careful reading of the paper and their 
suggestions. The issues and our replies are below.

Issue 1: • On page 1 the authors state ”the lightest of the MSSM particles known as the neutralinos emerges as the most attractive candidate for the LSP.”. It is not true that the lightest MSSM particles ought to be neutralinos (there is nothing wrong with sneutrinos being the LSP), and also I guess that one is not hunting for an attractive candidate for LSP, but rather trying to probe the particle nature of dark matter. I would suggest to correct this sentence accordingly.

\end{document}